---
title: "Getting started with jst"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with jst}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(jst)
```

## Overview

**jst** provides S7 classes that model every JSON value type. You get a typed
R object — not just a raw string — that serialises to valid JSON via
`format()` or `print()`.

| Constructor | JSON type | Example output |
|---|---|---|
| `json_null()` | `null` | `null` |
| `json_boolean(x)` | `true` / `false` | `true` |
| `json_number(x)` | number | `42` |
| `json_string(x)` | string | `"hello"` |
| `json_vector(x)` | typed array | `[1, 2, 3]` |
| `json_array(...)` | mixed array | `[1, "two", null]` |
| `json_object(...)` | object | `{"x": 1}` |

All types share the abstract `json` base class, so you can use `inherits(x,
"json")` to check for any JSON value.

## Scalar types

```{r scalars}
json_null()
json_boolean(TRUE)
json_boolean(FALSE)
json_number(42)
json_number(3.14)
json_string("hello, world")
```

Numbers are always serialised without scientific notation, and strings are
properly escaped:

```{r escaping}
json_number(1e6)          # not "1e+06"
json_string('say "hi"')   # internal quotes escaped
json_string("line1\nline2")
```

## Arrays

Use `json_vector()` for **homogeneous** arrays of numbers or strings — it
stores the raw R vector and is more efficient than `json_array()` for large
sequences:

```{r json-vector}
json_vector(1:10)
json_vector(c("apple", "banana", "cherry"))
json_vector(c(1.5, 2.5, 3.5))
```

Use `json_array()` for **mixed-type** arrays, including nested JSON values:

```{r json-array}
json_array(1, "two", TRUE, json_null())
json_array(json_boolean(FALSE), json_number(0), json_string(""))
```

Passing a plain R vector to `json_array()` expands it element-by-element:

```{r json-array-vector}
json_array(1:5)          # same as json_array(1, 2, 3, 4, 5)
```

## Objects

`json_object()` takes **named** arguments and produces a JSON object. Values
can be any R scalar, vector, or already-constructed JSON value:

```{r json-object}
json_object(name = "Alice", age = 30, active = TRUE)

json_object(
  id      = 1L,
  tags    = json_vector(c("r", "json")),
  address = json_object(city = "London", zip = "EC1A")
)
```

## Coercion with `to_json()`

`to_json()` converts plain R values to the most appropriate JSON type, so you
rarely have to call the low-level constructors by hand:

```{r to-json-scalars}
to_json(NULL)
to_json(TRUE)
to_json(99L)
to_json(3.14)
to_json("hello")
```

Vectors of length > 1 become arrays:

```{r to-json-vectors}
to_json(c(TRUE, FALSE, TRUE))   # json_array  (logical)
to_json(1:5)                    # json_vector (integer)
to_json(c("a", "b", "c"))       # json_vector (character)
```

Named lists become objects; unnamed lists become arrays:

```{r to-json-lists}
to_json(list(x = 1, y = "hi"))    # json_object
to_json(list(1, "two", TRUE))     # json_array
```

Already-constructed `json` values pass through unchanged:

```{r to-json-passthrough}
j <- json_boolean(TRUE)
identical(to_json(j), j)
```

## Building complex documents

Combine constructors and `to_json()` to build any JSON structure:

```{r complex}
doc <- json_object(
  version  = 1L,
  enabled  = TRUE,
  config   = json_object(
    retries = 3L,
    tags    = json_vector(c("prod", "v2"))
  ),
  items    = json_array(
    json_object(id = 1L, label = "first"),
    json_object(id = 2L, label = "second")
  ),
  metadata = json_null()
)

cat(format(doc))
```
